---
title: "Analysis_07082019"
author: "Rose Abramoff"
date: "7/8/2019"
output: html_document
---
#Timer and Toggles
```{r}
rm(list=ls())

master.start.time <- Sys.time()
```

#Load libraries
```{r}
 library(readxl)
 library(dplyr)
 library(ggplot2)
 library(raster)
 library(mice)
 library(VIM)
 library(corrplot)
 library(randomForest)
 library(forestFloor)
 library(rfUtilities)

figdir = "/Users/rzabramoff/Dropbox (Climate)/LSCE work files/Documents/Qmax/"
datadir <- "/Users/rzabramoff/Dropbox (Climate)/LSCE work files/Code/DOC_potential/"
```

##Make plot of data locations
```{r}
makelabelsEW <- function(x) {ifelse(x < 0, parse(text=paste0(x,"^o", "*W")), ifelse(x > 0, parse(text=paste0(x,"^o", "*E")),x))}
makelabelsNS <- function(x) {ifelse(x < 0, parse(text=paste0(x,"^o", "*S")), ifelse(x > 0, parse(text=paste0(x,"^o", "*N")),x))}

xbreaks <- seq(-180,180,60)
xlabels <- makelabelsEW(xbreaks)

ybreaks <- seq(-90,90,30)
ylabels <- makelabelsNS(ybreaks)

wrld <- map_data("world", ylim=c(-60,90))

gg1 <- ggplot() + 
  geom_polygon(data = wrld, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  coord_fixed(1.3) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), legend.position="bottom",legend.title=element_text(size=15), 
    legend.text=element_text(size=12), axis.text = element_text(size=12))+
 scale_x_continuous("", breaks = xbreaks, labels = xlabels) +
 scale_y_continuous("", breaks = ybreaks, labels = ylabels)

barsize=10
```

##Load multiplot function
```{r}
mypal <- c("#CC79A7", "#009E73", "#0072B2", "#D55E00", "#56B4E9","#E69F00", "#F0E442","#999999")

multiplot <- function(..., plotlist=NULL, file, cols=2, layout=NULL) {
  library(grid)
  
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

#Prep relationships
##Read in dataset
```{r}
print.df <- read.csv(file=paste0(datadir,"printdf.csv"))
```

##Add climate covariates
```{r}
load(file=paste0(datadir,"climatevars.Rdata"))

map.points <- cbind(print.df$Longitude, print.df$Latitude)
print.df$MAT.C.map <- raster::extract(mat, map.points)
print.df$MAP.mm.map <- raster::extract(map, map.points)

plot(print.df$MAT.C, print.df$MAT.C.map)
lines(c(0,60000),c(0,60000), lty=2)

plot(print.df$MAP.mm, print.df$MAP.mm.map)
lines(c(0,60000),c(0,60000), lty=2)

summary(lm(print.df$MAT.C.map ~ print.df$MAT.C))
summary(lm(print.df$MAP.mm.map ~ print.df$MAP.mm))
```

##Make map of locations
```{r}
pdf(file=paste0(figdir,"Map_of_locations.pdf"), height=4, width=7)
gg1 + geom_point(data=print.df, aes(x=Longitude, y=Latitude), color="red", pch=1, size=1)
dev.off()
```

##Impute missing data
```{r}
#Subset of larger dataset to contain only relevant columns for analyses
depth.key <- read_excel(paste0(datadir,"depth_key.xlsx"))

temp.df <- left_join(print.df, depth.key, by="Depth.cm")

missing.df <- temp.df[,c("Qmax.mg.kg","k.L.mg","Clay.per","Fed.g.kg","pH","OC.mg.kg","Alo.g.kg","Ald.g.kg","Alp.g.kg","Feo.g.kg","Fep.g.kg","Caex.cmol.kg","Silt.per","Sand.per","Soil.Order","Mid.Depth.cm","BD.g.cm3","MAT.C.map","MAP.mm.map")]
missing.df <- missing.df %>%
  dplyr::rename(MAT.C = MAT.C.map, MAP.mm = MAP.mm.map)

#Interpolate missing data
out <- mice(missing.df, seed=1)
bigdf <- mice::complete(out)
bigdf$Dataset <- print.df$Dataset

pdf(file=paste0(figdir,"FigureS4.pdf"), height=8, width=8)
densityplot(out, subset=.imp==0:1)
dev.off()

#Proportion missing
#Silt and Sand are the most missing
all <- dim(missing.df)[1]
maxmiss <- dim(missing.df[!is.na(missing.df$Silt.per),])[1]
(all-maxmiss)/all

#How much data are missing for pH and Clay.per?
(all-dim(missing.df[!is.na(missing.df$pH),])[1])/all
(all-dim(missing.df[!is.na(missing.df$Clay.per),])[1])/all
```

##Visualize Imputed Data
```{r}
pdf(file=paste0(figdir,"FigureS2.pdf"), height=3, width=7)
aggr = aggr(missing.df, col=mdc(1:2), numbers=TRUE, sortVars=TRUE, labels=names(missing.df), cex.axis=.7, gap=3, ylab=c("Proportion of missingness","Missingness Pattern"))
dev.off()

pairs.df <- missing.df[,c("Qmax.mg.kg", "k.L.mg", "Clay.per", "Fed.g.kg", "pH", "OC.mg.kg", "Alo.g.kg", "Ald.g.kg", "Alp.g.kg", "Feo.g.kg", "Fep.g.kg", "Caex.cmol.kg", "Silt.per", "Sand.per","MAT.C","MAP.mm")]

dfcor <- cor(pairs.df, use="complete.obs")

pdf(file=paste0(figdir,"FigureS3.pdf"), width=11, height=8)
corrplot(dfcor, method="shade", shade.col=NA, tl.col="black", tl.srt=45, cl.cex=1.5, tl.cex = 1.5)
dev.off()
```

##Random Forest
```{r}
q.df <- bigdf[,c("Qmax.mg.kg","Clay.per","Fed.g.kg","pH","OC.mg.kg","Alo.g.kg","Ald.g.kg","Alp.g.kg","Feo.g.kg","Fep.g.kg","Caex.cmol.kg","Silt.per","Sand.per","Soil.Order","Mid.Depth.cm","BD.g.cm3","MAT.C","MAP.mm")] 
k.df <- bigdf[,c("k.L.mg","Clay.per","Fed.g.kg","pH","OC.mg.kg","Alo.g.kg","Ald.g.kg","Alp.g.kg","Feo.g.kg","Fep.g.kg","Caex.cmol.kg","Silt.per","Sand.per","Soil.Order","Mid.Depth.cm","BD.g.cm3","MAT.C","MAP.mm")] 
q.names = k.names = c("Clay %", expression(Fe[d]~"(g/kg)"), "pH", "OC (mg/kg)", expression(Al[o]~"(g/kg)"), expression(Al[d]~"(g/kg)"),
expression(Al[p]~"(g/kg)"),
expression(Fe[o]~"(g/kg)"),
expression(Fe[p]~"(g/kg)"),
expression(Ca[ex]~"(cmol/kg)"),
"Silt %", "Sand %", "Soil Order", "Depth (cm)", "BD (g/cm3)","MAT (C)","MAP (mm)")

q.names = c(expression(Al[d]~"(g/kg)"),
            "Silt %", 
            "Sand %",
            "Depth (cm)",
            expression(Fe[p]~"(g/kg)"),
            expression(Al[o]~"(g/kg)"),
            "BD (g/cm3)",
            "pH",
            expression(Ca[ex]~"(cmol/kg)"),
            "OC (mg/kg)", 
            expression(Al[p]~"(g/kg)"),
            expression(Fe[o]~"(g/kg)"),
            "MAP (mm)",
            "Soil Order", 
            "Clay %",
            "MAT (C)",
            expression(Fe[d]~"(g/kg)")  
            )
k.names = c("Silt %",
            "Sand %",
            "Depth (cm)", 
            "BD (g/cm3)",
            expression(Ca[ex]~"(cmol/kg)"),
            expression(Al[o]~"(g/kg)"),
            expression(Al[d]~"(g/kg)"),
            expression(Fe[p]~"(g/kg)"),
            expression(Al[p]~"(g/kg)"),
            expression(Fe[o]~"(g/kg)"),
            expression(Fe[d]~"(g/kg)"),
            "Clay %",
            "MAP (mm)",
            "Soil Order", 
            "MAT (C)",
            "pH",
            "OC (mg/kg)"
            )
#log transform y-variables
q.df$Qmax.mg.kg <- log(q.df$Qmax.mg.kg)
k.df$k.L.mg <- log(k.df$k.L.mg)

#clean missing values
q.df <- q.df[complete.cases(q.df),]
k.df <- k.df[complete.cases(k.df),]

q.df.long <- q.df
k.df.long <- k.df

set.seed(1)
q.rf <- randomForest(Qmax.mg.kg ~ ., data=q.df, importance=T, proximity=T, type = "regression") 
set.seed(1)
k.rf <- randomForest(k.L.mg ~ ., data=k.df, importance=T, proximity=T, type = "regression")
print(paste(q.rf$mse[q.rf$ntree],"qmax MSE"))
print(paste(q.rf$rsq[q.rf$ntree],"qmax R2"))
print(paste(k.rf$mse[k.rf$ntree],"kaff MSE"))
print(paste(k.rf$rsq[k.rf$ntree],"kaff R2"))

#generate variable importance (mean decrease in node impurity/mean increase in node purity)
q.imp = round(importance(q.rf, type=2, scale=T),2)[,1]
k.imp = round(importance(k.rf, type=2, scale=T),2)[,1]

pdf(file=paste0(figdir,"Figure2.pdf"), width=7, height=4.25)
par(mfrow=c(1,2))
par(mar = c(3.5,6,1,0.5)+.3)
barplot(sort(q.imp), horiz=TRUE, las=2, main=expression(Q[sp]~ "(mg/kg)"), names.arg=q.names)
barplot(sort(k.imp), horiz=TRUE, las=2, main=expression(k~ "(L/mg)"), names.arg=k.names)
dev.off()
```

###Save an RF model
```{r}
q.df <- bigdf[,c("Qmax.mg.kg","Clay.per","pH","Soil.Order","MAT.C","MAP.mm")]
k.df <- bigdf[,c("k.L.mg","Clay.per","pH","Soil.Order","MAT.C","MAP.mm")] 
q.names = k.names <- c("Clay %", "pH", "Soil.Order","MAT.C","MAP.mm")

#log transform y-variables
q.df$Qmax.mg.kg <- log(q.df$Qmax.mg.kg)
k.df$k.L.mg <- log(k.df$k.L.mg)

#clean missing values
q.df <- q.df[complete.cases(q.df),]
k.df <- k.df[complete.cases(k.df),]

set.seed(1)
q.rf <- randomForest(Qmax.mg.kg ~ ., data=q.df, importance=T, proximity=T, type = "regression", keep.inbag=T) 
set.seed(1)
k.rf <- randomForest(k.L.mg ~ ., data=k.df, importance=T, proximity=T, type = "regression", keep.inbag=T)

print(paste(q.rf$mse[q.rf$ntree],"qmax MSE"))
print(paste(q.rf$rsq[q.rf$ntree],"qmax R2"))
print(paste(k.rf$mse[k.rf$ntree],"kaff MSE"))
print(paste(k.rf$rsq[k.rf$ntree],"kaff R2"))

#generate variable importance (mean decrease in node impurity/mean increase in node purity)
q.imp = round(importance(q.rf, type=2, scale=T),2)[,1]
k.imp = round(importance(k.rf, type=2, scale=T),2)[,1]

par(mfrow=c(1,2))
par(mar = c(3.5,6,1,0.5)+.3)
barplot(q.imp, horiz=TRUE, las=2, main=expression(Q[sp]~ "(mg/kg)"), names.arg=q.names)
barplot(k.imp, horiz=TRUE, las=2, main=expression(k~ "(L/mg)"), names.arg=k.names)

save(q.rf, k.rf, file=paste0(datadir,"/bestRFmodels.Rdata"))
```

###RF Cross-validation Methods
```{r}
q.rf.cv <- rf.crossValidation(q.rf, q.df[,!(names(q.df) %in% "Qmax.mg.kg")] , p = 0.2, n = 99, seed = 1)
q.rf.cv$fit.var.exp
q.rf.cv$fit.mse

k.rf.cv <- rf.crossValidation(k.rf, k.df[,!(names(k.df) %in% "k.L.mg")] , p = 0.2, n = 99, seed = 1)
k.rf.cv$fit.var.exp
k.rf.cv$fit.mse
```

###RF CV generate new models
```{r}
q.df <- bigdf[,c("Qmax.mg.kg","Clay.per","pH","Soil.Order","MAT.C","MAP.mm")]
q.df$Qmax.mg.kg <- log(q.df$Qmax.mg.kg)
q.df <- q.df[complete.cases(q.df),]

set.seed(1)
ss <- sample(rep(1:5, diff(floor(nrow(q.df) * c(0, 0.2, 0.4, 0.6, 0.8, 1)))))
random_5 <- setNames(split(q.df,ss), c("1","2","3","4","5"))
q.df$random5 <- ss

set.seed(1)
q.df.sub <- q.df[q.df$random5!="1",-7]
q.rf1 <- randomForest(Qmax.mg.kg ~ ., data=q.df.sub, importance=T, proximity=T, type = "regression", keep.inbag=T) 
set.seed(1)
q.df.sub <- q.df[q.df$random5!="2",-7]
q.rf2 <- randomForest(Qmax.mg.kg ~ ., data=q.df.sub, importance=T, proximity=T, type = "regression", keep.inbag=T) 
set.seed(1)
q.df.sub <- q.df[q.df$random5!="3",-7]
q.rf3 <- randomForest(Qmax.mg.kg ~ ., data=q.df.sub, importance=T, proximity=T, type = "regression", keep.inbag=T) 
set.seed(1)
q.df.sub <- q.df[q.df$random5!="4",-7]
q.rf4 <- randomForest(Qmax.mg.kg ~ ., data=q.df.sub, importance=T, proximity=T, type = "regression", keep.inbag=T) 
set.seed(1)
q.df.sub <- q.df[q.df$random5!="5",-7]
q.rf5 <- randomForest(Qmax.mg.kg ~ ., data=q.df.sub, importance=T, proximity=T, type = "regression", keep.inbag=T) 

save(q.rf1, q.rf2, q.rf3, q.rf4, q.rf5, file=paste0(datadir,"CVRFmodels.Rdata"))
```

###Do partial dependence
```{r}
partialq <- forestFloor(rf.fit = q.rf,
                        X = q.df,
                        calc_np = F,
                        binary_reg = F,
                        impType=2)

partialk <- forestFloor(rf.fit = k.rf,
                        X = k.df,
                        calc_np = F,
                        binary_reg = F,
                        impType=2)

```
####Make partial plots
```{r}
#plots cross-validated partial feature contributions (y-axis) depending on the variable value (x-axis)

x_line <- 2.5
y_line <- 2
x_cex <- 0.85
y_cex <- 0.7
y_at_q <- -0.2
y_at_k <- -0.4

#prepare color legend
breaks.clay = c(1, 2, 3, 4, 5, 6, 7, 8)
labels.clay = c(0, 10, 20, 30, 40, 50, 60, 70)
breaks.pH = c(1, 2.4, 3.8, 5.2, 6.6, 8)
labels.pH = c(3, 4, 5, 6, 7, 8)
ColRange = c("#FF000AB3","#D90600B3","#B34300B3","#6F8F00B3","#009274B3","#004BB6B3","#0014DCB3","#000AFFB3")

dummy <-  as.data.frame(cbind(x = 1:402, y=1:402, z= seq(1,8,length=402)))

pdf(file=paste0(figdir, "legend_clay.pdf"), height =2, width =2)
ggplot(dummy, aes(x,y, fill=z)) + geom_raster() + scale_fill_gradientn(colours=ColRange, breaks=breaks.clay,labels=labels.clay, name= "Clay (%)")
dev.off()

pdf(file=paste0(figdir, "legend_pH.pdf"), height =2, width =2)
ggplot(dummy, aes(x,y, fill=z)) + geom_raster() + scale_fill_gradientn(colours=ColRange, breaks=breaks.pH,labels=labels.pH, name= "pH")
dev.off()

pdf(file=paste0(figdir,"Figure3a.pdf"), height=4, width=6)
Col=fcol(partialq,1,orderByImportance=T)
par(mfrow=(c(2,3)))
par(mar=c(3.5,3.5,0.5,0), pch=16)
plot(partialq$X[1]$Clay.per, partialq$FCmatrix[,1],col=Col, ylab="", xlab="", ylim=c(-0.8,0.5)) 
mtext(text = expression(bold("Clay (%)")), side = 1, line=x_line, cex=x_cex)
mtext(text = expression(bold("Q"[sp]~" Partial Feature Contribution")), side = 2, line=y_line, cex=y_cex, at=y_at_q)
plot(partialq$X[2]$pH, partialq$FCmatrix[,2],col=Col, xlab="", ylab="", ylim=c(-0.8,0.5)) 
mtext(text = expression(bold("pH")), side = 1, line=x_line, cex=x_cex)
plot(partialq$X[3]$Soil.Order, partialq$FCmatrix[,3],col="darkgray", xlab="", ylab="", ylim=c(-0.8,0.5), xaxt="n")
text(seq(1, 6, by=1), par("usr")[3]-0.075, labels = c("Alfisol","Entisol","Inceptisol","Mollisol","Spodosol","Ultisol"), srt = 30, pos = 1, xpd = TRUE)
mtext(text = expression(bold("Soil Order")), side = 1, line=x_line, cex=x_cex)
plot(partialq$X[4]$MAT.C, partialq$FCmatrix[,4],col=Col, xlab="", ylab="", ylim=c(-0.8,0.5)) 
mtext(text = expression(bold("MAT ("~degree~"C)")), side = 1, line=x_line, cex=x_cex)
mtext(text = expression(bold("Q"[sp]~" Partial Feature Contribution")), side = 2, line=y_line, cex=y_cex, at=y_at_q)
plot(partialq$X[5]$MAP.mm, partialq$FCmatrix[,5],col=Col, pch=16, xlab="", ylab="", ylim=c(-0.8,0.5))
mtext(text = expression(bold("MAP (mm)")), side = 1, line=x_line, cex=x_cex)
dev.off()

pdf(file=paste0(figdir,"Figure3b.pdf"), height=4, width=6)
Col=fcol(partialk,1,orderByImportance=T)
par(mfrow=(c(2,3)))
par(mar=c(3.5,3.5,0.5,0), pch=16)
plot(partialk$X[1]$Clay.per, partialk$FCmatrix[,1],col=Col, xlab="", ylab="", ylim=c(-1.2,0.6)) 
mtext(text = expression(bold("Clay (%)")), side = 1, line=x_line, cex=x_cex)
mtext(text = expression(bold("k Partial Feature Contribution")), side = 2, line=y_line, cex=y_cex, at=y_at_k)
plot(partialk$X[2]$pH, partialk$FCmatrix[,2],col=Col, xlab= "", ylab="", ylim=c(-1.2,0.6)) 
mtext(text = expression(bold("pH")), side = 1, line=x_line, cex=x_cex)
plot(partialk$X[3]$Soil.Order, partialk$FCmatrix[,3],col="darkgray", xlab="", ylab="", ylim=c(-1.2,0.6), xaxt="n")
text(seq(1, 6, by=1), par("usr")[3]-0.1, labels = c("Alfisol","Entisol","Inceptisol","Mollisol","Spodosol","Ultisol"), srt = 30, pos = 1, xpd = TRUE)
mtext(text = expression(bold("Soil Order")), side = 1, line=x_line, cex=x_cex)
plot(partialk$X[4]$MAT.C, partialk$FCmatrix[,4],col=Col, xlab="", ylab="", ylim=c(-1.2,0.6)) 
mtext(text = expression(bold("MAT ("~degree~"C)")), side = 1, line=x_line, cex=x_cex)
mtext(text = expression(bold("k Partial Feature Contribution")), side = 2, line=y_line, cex=y_cex, at=y_at_k)
plot(partialk$X[5]$MAP.mm, partialk$FCmatrix[,5],col=Col, pch=16, xlab="", ylab="", ylim=c(-1.2,0.6)) 
mtext(text = expression(bold("MAP (mm)")), side = 1, line=x_line, cex=x_cex)
dev.off()
```

####Save ranges of predictor variables in dataset
```{r}
dlims <- matrix(0,4,2)
dlims[1,1:2] <- range(bigdf$Clay.per)
dlims[2,1:2] <- range(bigdf$pH)
dlims[3,1:2] <- range(bigdf$MAT.C)
dlims[4,1:2] <- range(bigdf$MAP.mm)

rownames(dlims) <- c("Clay.per","pH","MAT.C","MAP.mm")
```

#Extrapolate Qmax and K to WoSIS profiles
##Hex Plot Settings
```{r}
load(file=paste0(datadir,"wosis.Rdata"))
load(file=paste0(datadir,"climatevars.Rdata"))

color.setting.K = col=rev(terrain.colors(255))
color.setting.Q = col=rev(terrain.colors(255))
```

##Separate topsoils 0-30cm with subsoils 30cm-1m
```{r}
load(file=paste0(datadir,"indexdf.Rdata"))
lay$top.num <- as.numeric(as.character(lay$top))
lay$bottom.num <- as.numeric(as.character(lay$bottom))

topsoil <- lay[lay$bottom.num<30,]
subsoil <- lay[lay$bottom.num>30 & lay$bottom.num<100,]

temp.topsoil <- merge(topsoil, indexdf, by="profile_id")
temp.subsoil <- merge(subsoil, indexdf, by="profile_id")
topsoil <- temp.topsoil[temp.topsoil$same.clay==F && temp.topsoil$same.ph==F,]
subsoil <- temp.subsoil[temp.subsoil$same.clay==F && temp.subsoil$same.ph==F,]
```

##Estimating saturation characteristics
###Compute profiles function with random forest
```{r}
#BDWS #bulk density whole soil (kg/dm3)
#CLAY #clay total (g/100g)
#PHCA #pH CaCl2 (same as Mayes et al. 2012)
#ORGC #organic carbon (g/kg)

format_lay <- function(lay) {
  
newdf <- lay %>%
        dplyr::group_by(profile_id)%>% 
        dplyr::summarise(BD.g.cm3 = mean(bdws_value_avg, na.rm=T),
                  Clay.per = mean(clay_value_avg, na.rm=T),
                  pH = mean(phca_value_avg, na.rm=T),
                  OC.g.kg = mean(orgc_value_avg, na.rm=T),
                  Coarse.per = mean(cfvo_value_avg, na.rm=T)
                  )

#remove pH outliers
newdf$pH[newdf$pH>15] <- NaN

#join df and newdf
df$profile_id <- as.character(df$profile_id)
lldf <- left_join(newdf, df, by=c("profile_id"))

lldf <- lldf %>%
  dplyr::select(c(profile_id, BD.g.cm3, Clay.per, pH, OC.g.kg, cstx_order_name, MAT.C, MAP.mm, latitude, longitude, Coarse.per)) %>%
  dplyr::rename(Soil.Order = cstx_order_name)

#tot soc gC/kg soil * bulk density kg soil/dm3 * 1000dm3/m3 *1kg/1000g = kg/m3
lldf$tot.soc.kg.m3 <- lldf$OC.g.kg*lldf$BD.g.cm3
lldf$OC.mg.kg <- lldf$OC.g.kg*1000
#tempvalue.totsoc <- lldf$tot.soc.kg.m3

return(lldf)
}

compute_profiles <- function(lldf) {
lldf.complete <- lldf %>%
    tidyr::drop_na(Soil.Order, Clay.per, pH, MAT.C, MAP.mm) %>%
  dplyr::select(c(Clay.per, pH, Soil.Order, MAT.C, MAP.mm)) %>% 
  dplyr::filter(Soil.Order == "Alfisol" | Soil.Order == "Entisol" | Soil.Order == "Inceptisol" | Soil.Order == "Mollisol" | Soil.Order == "Spodosol" | Soil.Order == "Ultisol") %>%
  droplevels(exclude = "")

lldf.complete$k.L.mg <- exp(predict(k.rf, lldf.complete))

lldf.complete$qmax.mg.kg <- exp(predict(q.rf, lldf.complete))

lldf.complete$qmax.g.kg <- lldf.complete$qmax.mg.kg/1000

lldf <- dplyr::full_join(lldf.complete, lldf, by = c("Clay.per", "pH", "MAT.C", "MAP.mm"))

#after multiplying by BD, Qmax is in units g/m3
lldf$qmax.g.m3 = lldf$BD.g.cm3*lldf$qmax.mg.kg*(1-lldf$Coarse.per/100)

return(lldf)
}
```

###Compute profiles
```{r}
lldf <- format_lay(lay)
lldf.avg <- compute_profiles(lldf)
```

###Compute profiles separating topsoils and subsoils
```{r}
topsoil.new <- format_lay(topsoil)
subsoil.new <- format_lay(subsoil)

lldf.topsoil <- compute_profiles(topsoil.new)
lldf.subsoil <- compute_profiles(subsoil.new)
```

##Nice Hex Plots Function
```{r}
alphahex <- 0.8
make.hex.plots <- function(data.to.plot, nbins, mins, maxs){
pdf(file=paste0(figdir,"Figure4ad.pdf"), height=7, width=7)

ggQ <- gg1+
stat_summary_hex(data = data.to.plot, aes(x = longitude, y = latitude, z = qmax.g.kg), fun = function(z) sum(z)/length(z), alpha = alphahex, bins= nbins) +
scale_fill_gradientn(colours = color.setting.Q, name=expression("Q"[sp]~" (g/kg)"), limits = c(mins[1],maxs[1])) +
guides(fill = guide_colourbar(barwidth = barsize))

ggK <- gg1+ 
stat_summary_hex(data = data.to.plot, aes(x = longitude, y = latitude, z = k.L.mg), fun = function(z) sum(z)/length(z), alpha = alphahex, bins= nbins) +
scale_fill_gradientn(colours = color.setting.K, name="k (L/mg)", limits = c(mins[2],maxs[2]), breaks = c(0.05,0.1), labels = c("0.05","0.10")) +
guides(fill = guide_colourbar(barwidth = barsize))

print(ggQ)
print(ggK)
dev.off()
}

make.density.plots <- function(){
lldf.topsoil$ID <- rep("topsoil",dim(lldf.topsoil)[1])
lldf.subsoil$ID <- rep("subsoil",dim(lldf.subsoil)[1])
topsub <- bind_rows(lldf.topsoil,lldf.subsoil)

ggQ.dens <- ggplot(topsub, aes(x=qmax.g.kg, fill=ID)) +
  geom_density(alpha=0.3) +
  xlab(expression("Q"[sp]~" (g/kg)")) + ylab("Density") + theme_classic()+
  guides(fill = guide_legend(reverse = TRUE)) 

ggK.dens <- ggplot(topsub, aes(x=k.L.mg, fill=ID)) + 
  geom_density(alpha=0.3) +
  xlab("k (L/mg)") + ylab("Density") + theme_classic()+
  guides(fill = guide_legend(reverse = TRUE)) 

ggS.dens <- ggplot(topsub, aes(x=tot.soc.kg.m3, fill=ID)) + 
  geom_density(alpha=0.3) +
  xlab(expression("Total SOC (kg C " ~ m^{-3} ~ ")")) + ylab("Density") + theme_classic()

pdf(file=paste0(figdir,"FigureS7.pdf"),height=2.5,width=4)
print(ggQ.dens)
print(ggK.dens)
print(ggS.dens)
dev.off()
}
```

##Plot the plots
```{r}
lims <- read.csv(file=paste0(datadir,"limits.csv"))
mins <- lims$out.min
maxs <- lims$out.max

make.hex.plots(lldf.avg, 50, mins, maxs)
make.density.plots()
```

##Summaries
```{r}
quantile(lldf.avg$qmax.g.kg, na.rm=T, c(.025, .5, .975))
quantile(lldf.avg$k.L.mg, na.rm=T, c(.025, .5, .975))

##How many profiles have I measured protected C for?
length(lldf.avg$qmax.mg.kg[!is.na(lldf.avg$qmax.mg.kg)])
length(lldf.avg$k.L.mg[!is.na(lldf.avg$k.L.mg)])

length(lldf.topsoil$qmax.g.kg[!is.na(lldf.topsoil$qmax.g.kg)])
length(lldf.subsoil$qmax.g.kg[!is.na(lldf.subsoil$qmax.g.kg)])
length(lldf.topsoil$k.L.mg[!is.na(lldf.topsoil$k.L.mg)])
length(lldf.subsoil$k.L.mg[!is.na(lldf.subsoil$k.L.mg)])
topsub.labels <- c(rep("topsoil",dim(lldf.topsoil)[1]), rep("subsoil",dim(lldf.subsoil)[1]))
topsub.df <- cbind(rbind(lldf.topsoil,lldf.subsoil),topsub.labels)
t.test(topsub.df$qmax.g.kg ~ topsub.df$topsub.labels)

quantile(lldf.topsoil$qmax.g.kg, na.rm=T, c(.025, .5, .975))
quantile(lldf.subsoil$qmax.g.kg, na.rm=T, c(.025, .5, .975))

quantile(lldf.topsoil$k.L.mg, na.rm=T, c(.025, .5, .975))
quantile(lldf.subsoil$k.L.mg, na.rm=T, c(.025, .5, .975))
```
#Extrapolate Qmax and K to gridded maps
##Load Layers
```{r}
load(file=paste0(datadir,"maplayers.Rdata"))
resolution.num = 10
```

##Generate new map using relationships
###Organize data
```{r}
#Stacks
#Inputs: pH, Clay.per, BD.g.cm3, MAP.mm, MAT.C, Soil.Order
df.stack <- as.data.frame(values(weighted.phs)) #pH units
names(df.stack) <- c("pH")
df.stack$Clay.per <- values(weighted.clay) #percent clay
df.stack$BD.g.cm3 <- values(weighted.bd) #kg/dm3 or g/cm3
df.stack$tot.soc.kg.m3 <- values(converted.soc) #kg/m3
df.stack$MAT.C <- values(mat.sized) #Celsius
df.stack$MAP.mm <- values(map.sized) #mm
df.stack$Soil.Order.Int <- values(soilord) #integer value factor
df.stack$OC.mg.kg <- 1000*values(weighted.soc2) #g/kg
df.stack$Coarse.per <- values(weighted.cfrag) #percent volume coarse fragments
df.stack$MAT.C2 <- values(mat.sized2)
df.stack$MAP.mm2 <- values(map.sized2)
df.stack$Clay.per2 <- values(clay.sized2)
df.stack$pH2 <- values(phs.sized2)

int.key <- as.data.frame(soilord@data@attributes[[1]])
names(int.key) <- c("Soil.Order.Int","Soil.Order")

df.keep <- left_join(df.stack, int.key, by=c("Soil.Order.Int"))
df.keep$position <- 1:length(df.keep$pH)
df.next <- df.keep %>% 
  dplyr::filter(Soil.Order == "Alfisol" | Soil.Order == "Entisol" | Soil.Order == "Inceptisol" | Soil.Order == "Mollisol" | Soil.Order == "Spodosol" | Soil.Order == "Ultisol") %>%
droplevels(exclude = "") 
```

###Compute and plot stack functions
```{r}
compute_stack <- function(df.next) {
  
df.stack.complete <- df.next %>%
     tidyr::drop_na(Soil.Order.Int, Clay.per, BD.g.cm3, MAP.mm, pH, MAT.C, OC.mg.kg, Coarse.per)
 
k.L.mg.stack <- exp(predict(k.rf, df.stack.complete))
qmax.mg.kg.stack <- exp(predict(q.rf, df.stack.complete))

fit.result <- as.data.frame(cbind(df.stack.complete$position, k.L.mg.stack, qmax.mg.kg.stack))
names(fit.result) <- c("position","k.L.mg","qmax.mg.kg")
#expand back into position
df.scale <- left_join(df.keep, fit.result, by=c("position"))

df.scale$qmax.g.kg <- df.scale$qmax.mg.kg/1000

#after multiplying by BD, Qmax is in units g/m3
df.scale$qmax.g.m3 = df.scale$BD.g.cm3*df.scale$qmax.mg.kg*(1-df.scale$Coarse.per/100)
df.scale$qmax.kg.m3 = df.scale$qmax.g.m3/1000

return(df.scale)
 
}
 
make_plots <- function(df.scale, mins, maxs){
thing.to.plot.1 <- cbind(
  df.scale$qmax.g.kg,
  df.scale$k.L.mg, 
  df.scale$tot.soc.kg.m3)
plot.title.1 <- c(
  expression("Q"[sp]~" (g/kg)"), 
  "k (L/mg)", 
  expression("Total SOC (kg C " ~ m^{-3} ~ ")"))

for (i in 1:length(plot.title.1)){
#copy an old raster and replace with plotted values
plot.this <- weighted.phs
values(plot.this) <- thing.to.plot.1[,i]

plot.this.next <- as.data.frame(plot.this, xy=TRUE)
names(plot.this.next) <- c("long","lat","value")

if(i==2){
 print(ggplot()+
  geom_raster(data = plot.this.next, aes(x=long, y=lat, fill=value)) +
  geom_polygon(data = wrld, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  coord_fixed(1.3) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), legend.position="bottom",legend.title=element_text(size=15), 
    legend.text=element_text(size=12), axis.text = element_text(size=12))+
 scale_x_continuous("", breaks = xbreaks, labels = xlabels) +
 scale_y_continuous("", breaks = ybreaks, labels = ylabels) +
 scale_fill_gradientn(colours = rev(terrain.colors(255)), name=plot.title.1[i], na.value = NA, limits = c(mins[i], maxs[i]), breaks = c(0.05, 0.10), labels = c("0.05","0.10")) +
guides(fill = guide_colourbar(barwidth = barsize))) }else{
print(ggplot()+
  geom_raster(data = plot.this.next, aes(x=long, y=lat, fill=value)) +
  geom_polygon(data = wrld, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  coord_fixed(1.3) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), legend.position="bottom",legend.title=element_text(size=15), 
    legend.text=element_text(size=12), axis.text = element_text(size=12))+
 scale_x_continuous("", breaks = xbreaks, labels = xlabels) +
 scale_y_continuous("", breaks = ybreaks, labels = ylabels) +
 scale_fill_gradientn(colours = rev(terrain.colors(255)), name=plot.title.1[i], na.value = NA, limits = c(mins[i], maxs[i])) +
guides(fill = guide_colourbar(barwidth = barsize)))
}
}

}
```

###Compute stack for uncertainty
```{r}
compute_stack_uncertainty <- function(df.stack.complete,q.rf.sub) {
 
k.L.mg.stack <- exp(predict(k.rf, df.stack.complete))
qmax.mg.kg.stack <- exp(predict(q.rf.sub, df.stack.complete))

fit.result <- as.data.frame(cbind(df.stack.complete$position, k.L.mg.stack, qmax.mg.kg.stack, df.stack.complete$Clay.per))
names(fit.result) <- c("position","k.L.mg","qmax.mg.kg","newClay.per")
#expand back into position
df.scale <- left_join(df.keep, fit.result, by=c("position"))

df.scale$qmax.g.kg <- df.scale$qmax.mg.kg/1000

#after multiplying by BD, Qmax is in units g/m3
df.scale$qmax.g.m3 = df.scale$BD.g.cm3*df.scale$qmax.mg.kg*(1-df.scale$Coarse.per/100)
df.scale$qmax.kg.m3 = df.scale$qmax.g.m3/1000

glob_qmax_for_uncert <- 1e-12*sum(df.scale$qmax.kg.m3*1e6*resolution.num^2,na.rm=T) 

return(glob_qmax_for_uncert)
 
}

df.stack.complete.repClay <- df.next %>%
    tidyr::drop_na(Soil.Order.Int, Clay.per2, BD.g.cm3, MAP.mm, pH, MAT.C, OC.mg.kg, Coarse.per) %>%
    dplyr::select(-c(Clay.per)) %>% 
    dplyr::rename(Clay.per = Clay.per2)

df.stack.complete.reppH <- df.next %>%
    tidyr::drop_na(Soil.Order.Int, Clay.per, BD.g.cm3, MAP.mm, pH2, MAT.C, OC.mg.kg, Coarse.per) %>%
    dplyr::select(-c(pH)) %>% 
    dplyr::rename(pH = pH2)

df.stack.complete.repMAT <- df.next %>%
    tidyr::drop_na(Soil.Order.Int, Clay.per, BD.g.cm3, MAP.mm, pH, MAT.C2, OC.mg.kg, Coarse.per) %>%
    dplyr::select(-c(MAT.C)) %>% 
    dplyr::rename(MAT.C = MAT.C2)

df.stack.complete.repMAP <- df.next %>%
    tidyr::drop_na(Soil.Order.Int, Clay.per, BD.g.cm3, MAP.mm2, pH, MAT.C, OC.mg.kg, Coarse.per) %>%
    dplyr::select(-c(MAP.mm)) %>% 
    dplyr::rename(MAP.mm = MAP.mm2)

uncert <- NULL
uncert[1] <- compute_stack_uncertainty(df.stack.complete.repClay,q.rf1)
uncert[2] <- compute_stack_uncertainty(df.stack.complete.reppH,q.rf1)
uncert[3] <- compute_stack_uncertainty(df.stack.complete.repMAT,q.rf1)
uncert[4] <- compute_stack_uncertainty(df.stack.complete.repMAP,q.rf1)

uncert[5] <- compute_stack_uncertainty(df.stack.complete.repClay,q.rf2)
uncert[6] <- compute_stack_uncertainty(df.stack.complete.reppH,q.rf2)
uncert[7] <- compute_stack_uncertainty(df.stack.complete.repMAT,q.rf2)
uncert[8] <- compute_stack_uncertainty(df.stack.complete.repMAP,q.rf2)

uncert[9] <- compute_stack_uncertainty(df.stack.complete.repClay,q.rf3)
uncert[10] <- compute_stack_uncertainty(df.stack.complete.reppH,q.rf3)
uncert[11] <- compute_stack_uncertainty(df.stack.complete.repMAT,q.rf3)
uncert[12] <- compute_stack_uncertainty(df.stack.complete.repMAP,q.rf3)

uncert[13] <- compute_stack_uncertainty(df.stack.complete.repClay,q.rf4)
uncert[14] <- compute_stack_uncertainty(df.stack.complete.reppH,q.rf4)
uncert[15] <- compute_stack_uncertainty(df.stack.complete.repMAT,q.rf4)
uncert[16] <- compute_stack_uncertainty(df.stack.complete.repMAP,q.rf4)

uncert[17] <- compute_stack_uncertainty(df.stack.complete.repClay,q.rf5)
uncert[18] <- compute_stack_uncertainty(df.stack.complete.reppH,q.rf5)
uncert[19] <- compute_stack_uncertainty(df.stack.complete.repMAT,q.rf5)
uncert[20] <- compute_stack_uncertainty(df.stack.complete.repMAP,q.rf5)

sd(uncert)
```

###Make plots of the input layers
```{r}
make_input_plots <- function(df.scale, mins, maxs){
thing.to.plot.1 <- cbind(
  df.scale$Clay.per,
  df.scale$pH, 
  df.scale$Soil.Order, 
  df.scale$MAT.C, 
  df.scale$MAP.mm)
plot.title.1 <- c(
  "Clay (%)", 
  "pH", 
  "Soil Order",
  "MAT (C)", 
  "MAP (mm)")

for (i in 1:length(plot.title.1)){
#copy a raster template and replace with plotted values
plot.this <- weighted.phs
values(plot.this) <- thing.to.plot.1[,i]

plot.this.next <- as.data.frame(plot.this, xy=TRUE)
names(plot.this.next) <- c("long","lat","value")

print(ggplot()+
  geom_raster(data = plot.this.next, aes(x=long, y=lat, fill=value)) +
  geom_polygon(data = wrld, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  coord_fixed(1.3) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), legend.position="bottom",legend.title=element_text(size=15), 
    legend.text=element_text(size=12), axis.text = element_text(size=12))+
 scale_x_continuous("", breaks = xbreaks, labels = xlabels) +
 scale_y_continuous("", breaks = ybreaks, labels = ylabels) +
 scale_fill_gradientn(colours = rev(terrain.colors(255)), name=plot.title.1[i], na.value = NA, limits = c(mins[i], maxs[i])) +
guides(fill = guide_colourbar(barwidth = barsize)))
}

}
```

##Analysis
```{r}
df.scale <- compute_stack(df.next)

pdf(file=paste0(figdir,"Figure4be.pdf"), height=7, width=7)
make_plots(df.scale, mins, maxs)
dev.off()

pdf(file=paste0(figdir,"FigureS5abcd.pdf"), height=7, width=7)
make_input_plots(df.scale, c(0,3.565,0,-27.57726,0), c(71.575,9.260,12,33.88745,8338.148))
dev.off()

plot.this <- weighted.phs
values(plot.this) <- df.scale$Soil.Order

temp.grid <- as.data.frame(plot.this, xy=TRUE)
names(temp.grid) <- c("long","lat","value")

pdf(file=paste0(figdir,"FigureS5e.pdf"), height=7, width=7)
ggplot()+
  geom_raster(data = temp.grid, aes(x=long, y=lat, fill=value)) +
  geom_polygon(data = wrld, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  coord_fixed(1.3) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), legend.position="bottom",legend.title=element_text(size=15), 
    legend.text=element_text(size=12), axis.text = element_text(size=12))+
 scale_x_continuous("", breaks = xbreaks, labels = xlabels) +
 scale_y_continuous("", breaks = ybreaks, labels = ylabels) +
 scale_fill_discrete(name="Soil Order", na.value = NA)
dev.off()

plot.this <- weighted.phs
values(plot.this) <- df.scale$qmax.g.kg
plot.this.next <- as.data.frame(plot.this, xy=TRUE)
names(plot.this.next) <- c("long","lat","value")

lat.plot <- plot.this.next %>%
  group_by(lat) %>% 
  dplyr::summarise(q.med = median(value, na.rm=T), q.5 = quantile(value, na.rm=T, .025), q.95 = quantile(value, na.rm=T, .975))

pdf(file=paste0(figdir,"Figure4c.pdf"), height=4, width=3)
print(ggplot()+
  geom_rect(data = lat.plot, aes(ymin=q.5, ymax=q.95, xmin=lat, xmax=lat+0.1)) +
  geom_line(data = lat.plot, aes(y=q.med, x = lat), color = "black") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_blank(), axis.text = element_text(size=12))+
 scale_x_continuous("", breaks = ybreaks, labels = ylabels) +
  ylab(expression(Q[sp]~ "(g/kg)")) +
  coord_flip())
dev.off()
```

##Summaries
```{r}
get_global_summaries <- function(df.scale) {
#get global qmax estimate
glob_qmax <- 1e-12*sum(df.scale$qmax.kg.m3*1e6*resolution.num^2,na.rm=T) #convert kg/m2 to 1m to kg/10km2 (1000m*1000m*10*10/km2) #kg to gigatonnes

#get total soc in the 6 soil orders
df.scale.drops <- df.scale %>%
    tidyr::drop_na(qmax.kg.m3)
glob_total_SOC_in_6_orders = sum(df.scale.drops$tot.soc.kg.m3*1e6*resolution.num^2,na.rm=T)*1e-12

outsummary = c(glob_qmax, glob_total_SOC_in_6_orders)
names(outsummary) =  c("glob_qmax", "glob_total_SOC_in_6_orders")
return(outsummary)

}

get_global_summaries(df.scale)
```

#Calculate shared min and max limits
```{r,eval=F}
#names
names <- c("qmax","kaff","totalSOC")

#plot range summaries for lldf.avg
point <- cbind(lldf.avg$qmax.g.kg,
               lldf.avg$k.L.mg,
               lldf.avg$tot.soc.kg.m3
               )

#plot range summaries for df.scale
map <- cbind(df.scale$qmax.g.kg,
             df.scale$k.L.mg,
             df.scale$tot.soc.kg.m3
             )

out.min = out.max = NULL

for (i in 1:ncol(point)){
out.min[i] <- min(range(point[,i], na.rm=T)[1],range(map[,i], na.rm=T)[1])
out.max[i] <- max(range(point[,i], na.rm=T)[2],range(map[,i], na.rm=T)[2])
}

lims <- as.data.frame(cbind(names, out.min, out.max))
lims$out.min <- as.numeric(as.character(lims$out.min))
lims$out.max <- as.numeric(as.character(lims$out.max))

write.csv(lims, file=paste0(datadir,"limits.csv"))
```

#Timer
```{r}
master.end.time <- Sys.time()
master.end.time - master.start.time
```

#Soil order plots and summaries
```{r}
quantile(bigdf$Qmax.mg.kg, na.rm=T, c(.25, .5, .75, .9))

pdf(file=paste0(figdir,"FigureS6ab.pdf"), width=6, height=3.5)
ggplot(bigdf, aes(x=bigdf$Soil.Order, y=bigdf$Qmax.mg.kg/1000)) + geom_violin() + geom_boxplot(width=0.1, fill="black") + stat_summary(fun.y=median, geom="point", fill="white", shape=21, size=2.5) +
xlab("Soil Order") +
ylab(expression(Q[sp]~ "(g/kg)"))

ggplot(bigdf, aes(x=bigdf$Soil.Order, y=bigdf$k.L.mg)) + geom_violin() + geom_boxplot(width=0.1, fill="black") + stat_summary(fun.y=median, geom="point", fill="white", shape=21, size=2.5) +
xlab("Soil Order") +
ylab(expression(k~ "(L/mg)"))

dev.off()

pdf(file=paste0(figdir,"FigureS6ab_inset.pdf"), width=4, height=1.5)

ggplot(bigdf, aes(x=bigdf$Soil.Order, y=bigdf$Qmax.mg.kg/1000)) + geom_violin() + geom_boxplot(width=0.1, fill="black") + stat_summary(fun.y=median, geom="point", fill="white", shape=21, size=2.5) +
xlab("Soil Order") +
ylab(expression(Q[sp]~ "(g/kg)")) +
ylim(0,5)

ggplot(bigdf, aes(x=bigdf$Soil.Order, y=bigdf$k.L.mg)) + geom_violin() + geom_boxplot(width=0.1, fill="black") + stat_summary(fun.y=median, geom="point", fill="white", shape=21, size=2.5) +
xlab("Soil Order") +
ylab(expression(k~ "(L/mg)")) +
ylim(0,0.3)

dev.off()
```

##Relationship between Qmax and TotalSOC
```{r}
fit <- summary(lm(df.scale$tot.soc.kg.m3~df.scale$qmax.g.m3))
sqrt(fit$r.squared)
```

